---
title: "ntimes: Kaplan Meier plots"
author: "Marcio A. Diniz"
format:
  html:
    theme: default
    html-table-processing: none
    toc: true
    toc-depth: 5
    toc-location: left
    column: page-right
    self-contained: true
    warning: false
    message: false
    fig-height: 5
    fig-width: 7
vignette: >
  %\VignetteIndexEntry{ntimes: Kaplan Meier plots}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}

---
# Libraries
```{r}
library(tidyverse)
library(survival)
library(kableExtra)
library(patchwork)

library(ntimes)
```

# Data
The dataset should be formatted to ensure that categorical and numerical variables have `factor` and `numeric` classes, respectively.

Furthermore, a list indicating the variable labels will be used in all plots.

```{r}
dataset <- lung |>
  mutate(sex = factor(sex, levels = 1:2,
                      labels = c("Female", "Male")),
         ph.ecog = as.factor(ph.ecog)) |>
  select(sex, ph.ecog, time, status)

data_labels <- list(sex = "Sex", ph.ecog = "ECOG")
```

# Kaplan-Meier plots

##  Standard Kaplan-Meier plots with labels

Standard Kaplan-Meier curves with labels can be obtained using the function `nt_km` with `time` and `status` as required inputs and `labels` as an optional input to format the plots. The output is a `list` with plots and tables for the overall survival and each variable in the data frame `dataset`.

```{r}
plots <- 
  dataset |> nt_km(time = time, status = status,
                   labels = data_labels,
                   base_size = 12)

```

::: {.panel-tabset}
### Overall
```{r}
plots$plot$overall$combined_plot

tab <- plots$tab$overall 
kable(tab) |> kable_classic()
```

### Sex
```{r}
plots$plot$sex$combined_plot

tab <- plots$tab$sex 
kable(tab) |> kable_classic()
```

### ECOG
```{r}
plots$plot$ph.ecog$combined_plot

tab <- plots$tab$ph.ecog
kable(tab) |> kable_classic()
```
:::

## Adding specific time points

A set of specific time points can be defined with the input `time_points` as a numeric vector. Furthermore, a table with the survival (95% CI) at the same specific time points can also be accessed.

```{r}
plots <- 
  dataset |> nt_km(time = time, status = status,
                   labels = data_labels,
                   time_points = c(200, 500, 800),
                   base_size = 12)

```

::: {.panel-tabset}
### Overall
```{r}
plots$plot$overall$combined_plot

tab <- plots$tab$overall 
kable(tab) |> kable_classic()
```

### Sex
```{r}
plots$plot$sex$combined_plot

tab <- plots$tab$sex 
kable(tab) |> kable_classic()
```

### ECOG
```{r}
plots$plot$ph.ecog$combined_plot

tab <- plots$tab$ph.ecog
kable(tab) |> kable_classic()
```
:::

## Changing x- and y-axis title labels

The x- and y-axis labels can be updated with `xlab` and `ylab`, respectively.

```{r}
plots <- 
  dataset |> nt_km(time = time, status = status,
                   labels = data_labels,
                   time_points = c(200, 500, 800),
                   xlab = "Time (Days)",
                   ylab = "Progression Free Survival",
                   base_size = 12)

```

::: {.panel-tabset}

### Overall
```{r}
plots$plot$overall$combined_plot
```

### Sex
```{r}
plots$plot$sex$combined_plot
```

### ECOG
```{r}
plots$plot$ph.ecog$combined_plot
```
:::


## Adding survival table

A table with median survival time (95% CI) or survival (95%) estimates at specific time points can also be added to Kaplan-Meier plots. A function should be provided as the input `survival_table` in the function `nt_km`. The function  `survival_table` is required to have two inputs: a `survfit` object and a label `var_label` for each covariate.

A standard `survival_table` can be obtained using the function `get_survival_table`. For example,

```{r}
my_survival_table <- get_survival_table()

fit <- survfit(Surv(time, status) ~ sex, data = dataset)
my_survival_table(fit)
```

Next, the Kaplan-Meier curves are plotted using the `my_survival_table` function as input for `survival_function`,

```{r}
plots <- 
  dataset |> nt_km(time = time, status = status,
                   labels = data_labels,
                   time_points = c(200, 500, 800),
                   survival_table = my_survival_table,
                   xlab = "Time (Days)",
                   ylab = "Progression Free Survival",
                   base_size = 12)
```

::: {.panel-tabset}
#### Overall
```{r}
plots$plot$overall$combined_plot
```

#### Sex
```{r}
plots$plot$sex$combined_plot
```

#### ECOG
```{r}
plots$plot$ph.ecog$combined_plot
```
:::

### Modifying survival table

There are two types (`type`) of information provided by the standard `survival_table` function from `get_survival_table`: `"median"` and `"time_points"`. If `type = "time_points"`, then the input `time_points` is required. Additionally, the `title`can be modified. 

First, the `survival_table` should carefully be tested:

```{r}
my_survival_table <- get_survival_table(title = "PFS, (95% CI)",
                                        type = "time_points",
                                        time_points = c(500))


fit <- survfit(Surv(time, status) ~ 1, data = dataset)
my_survival_table(fit)

fit <- survfit(Surv(time, status) ~ sex, data = dataset)
my_survival_table(fit)
```

Next, the Kaplan-Meier curves are plotted using the `survival_table` function,

```{r}
plots <- 
  dataset |> nt_km(time = time, status = status,
                   labels = data_labels,
                   time_points = c(200, 500, 800),
                   survival_table = my_survival_table,
                   xlab = "Time (Days)",
                   ylab = "Progression Free Survival")
```

::: {.panel-tabset}
#### Overall
```{r}
plots$plot$overall$combined_plot
```

#### Sex
```{r}
plots$plot$sex$combined_plot
```

#### ECOG
```{r}
plots$plot$ph.ecog$combined_plot
```
:::



### Adjusting figure size and font size

The font size can be adjusted using the argument `base_size` and figure height and width can be adjuste din Quarto using `fig-height` and `fig-width`.
```{r}
plots <- 
  dataset |> nt_km(time = time, status = status,
                   labels = data_labels,
                   time_points = c(200, 500, 800),
                   survival_table = my_survival_table,
                   base_size = 16,
                   xlab = "Time (Days)",
                   ylab = "Progression Free Survival")
```

::: {.panel-tabset}
#### Overall
```{r}
#| fig-height: 8
#| fig-width: 10
plots$plot$overall$combined_plot
```

#### Sex
```{r}
#| fig-height: 8
#| fig-width: 10
plots$plot$sex$combined_plot
```

#### ECOG
```{r}
#| fig-height: 8
#| fig-width: 10
plots$plot$ph.ecog$combined_plot
```
:::

## Further changes

Further changes can be done as the components of each plot are  `ggplot` objects.

```{r}
library(gridExtra)

plots$plot$ph.ecog$surv_plot <- 
  plots$plot$ph.ecog$surv_plot +
  theme_bw(base_size = 12)

custom_theme <- ttheme_default(
  core = list(fg_params = list(fontsize = 6,
                               col = "black"),
              bg_params = list(fill = "white",
                               col = "black", lwd = 1)),
  colhead = list(fg_params = list(fontsize = 6,
                                  fontface = "bold",
                                  col = "black"),
                 bg_params = list(fill = "white",
                                  col = "black", lwd = 1))
)

table_grob <- tableGrob(plots$plot$ph.ecog$surv_table,
                        rows = NULL,
                        theme = custom_theme)

plots$plot$ph.ecog$surv_plot <- 
  plots$plot$ph.ecog$surv_plot +
        annotation_custom(
          grob = table_grob,
          xmin = 600,
          xmax = 900,
          ymin = 0.7, ymax = 0.9
        )

plots$plot$ph.ecog$risk_table <- 
  plots$plot$ph.ecog$risk_table  +
  theme_bw(base_size = 12) +
  theme(axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

plots$plot$ph.ecog$combined_plot <-
  plots$plot$ph.ecog$surv_plot + 
  plots$plot$ph.ecog$risk_table + 
  plot_layout(nrow = 2, heights = c(0.8, 0.2))

plots$plot$ph.ecog$combined_plot
```

Extensive changes can be done editing the functions `std_km` and `std_km_group` using `edit(std_km)` or `edit(std_km_group)`, then creating new functions `my_std_km` and `my_std_km_group` to input as the arguments `std_km` and `std_km_group` in the function `nt_km`.


## Saving files
Finally, files can be saved in `jpeg` format

```{r, eval = FALSE}
plots <- 
  dataset |> nt_km(time = time, status = status,
                   labels = data_labels,
                   time_points = c(200, 500, 800),
                   survival_table = my_survival_table,
                   xlab = "Time (Days)",
                   ylab = "Progression Free Survival",
                   save = TRUE, 
                   fig_height = 8, fig_width = 10)
```

